---
title: "Vignette Title"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
knitr::opts_chunk$set(eval = FALSE)
```

# Installation and Setup 

R is used for most of the computations but there are external dependencies required for the packages needed to get this working.
Most of my this software requires a Linux/*nix machine.

If you're using a Docker/Singularity container, I suggest using Debian and Neurodebian (http://neuro.debian.net/).  

## External Dependencies
### R Package Setup

```{r, engine = "bash"}
sudo apt-get install r-base r-base-dev
```

In `R`:
```{r}
install.packages("devtools")
```

### FSL
FSL (https://fsl.fmrib.ox.ac.uk/fsl/fslwiki) must be installed for the fslr package (see https://github.com/muschellij2/fslr for additional Neurodebian setup).

After setting up the `apt-key`s sufficiently, the below may install FSL to work with `fslr` (not guaranteed).

```{r, engine = "bash"}
sudo apt-get install fsl-complete

FSLDIR=/usr/local/fsl
FSLSHARE=/usr/share/data

mkdir -p ${FSLDIR}/bin && cp /usr/lib/fsl/5.0/* ${FSLDIR}/bin/
mkdir -p ${FSLDIR}/data/standard && mkdir -p ${FSLDIR}/data/atlases 


#######################################
# Setting things up like other installers
#######################################
cp -R ${FSLSHARE}/fsl-mni152-templates/* ${FSLDIR}/data/standard/

# setting up atlases
cp -R ${FSLSHARE}/harvard-oxford-atlases/* ${FSLDIR}/data/atlases/ 
cp -R ${FSLSHARE}/juelich-histological-atlas/* ${FSLDIR}/data/atlases/ 
cp -R ${FSLSHARE}/bangor-cerebellar-atlas/* ${FSLDIR}/data/atlases/ 
cp -R ${FSLSHARE}/jhu-dti-whitematter-atlas/* ${FSLDIR}/data/atlases/ 
cp -R ${FSLSHARE}/forstmann-subthalamic-nucleus-atlas/* ${FSLDIR}/data/atlases/ 
cp -R ${FSLSHARE}/fsl-resting-connectivity-parcellation-atlases/* ${FSLDIR}/data/atlases/ 
cp -R ${FSLSHARE}/mni-structural-atlas/* ${FSLDIR}/data/atlases/ 
cp -R ${FSLSHARE}/oxford-thalamic-connectivity-atlas/* ${FSLDIR}/data/atlases/ 
cp -R ${FSLSHARE}/talairach-daemon-atlas/* ${FSLDIR}/data/atlases/ 

echo "export LD_LIBRARY_PATH=/usr/lib/fsl/5.0:$LD_LIBRARY_PATH" >> ~/.profile
echo "export LD_LIBRARY_PATH=/usr/lib/fsl/5.0:$LD_LIBRARY_PATH" >> ~/.bash_profile
```

### ANTsR and ITK-based software

The `ichseg` package relies upon `extrantsr`, which relies on `ANTsR`, `ANTsRCore`, and `ITKR`, which are largely powerful packages that rely upon the [ITK](https://itk.org) software.  

These rely on `git` and `cmake`, so they must be installed:

```{r, engine = "bash"}
sudo apt-get git-core
sudo apt-get cmake
```


In `R`:

```{r}
devtools::install_github("muschellij2/ITKR")
devtools::install_github("muschellij2/ANTsRCore")
devtools::install_github("muschellij2/ANTsR")
```


An easier way to install these packages is likely to use the binaries
#### OS X Binaries

The links for the OSX binaries are at:
```
https://github.com/muschellij2/ITKR/releases/download/v0.4.12.4/ITKR_0.4.12.4.tgz
https://github.com/muschellij2/ANTsRCore/releases/download/v0.4.2.1/ANTsRCore_0.4.2.1.tgz
https://github.com/muschellij2/ANTsR/releases/download/v0.6.2/ANTsR_0.6.2.tgz
```

#### Linux Binaries

The links for the Linux binaries are at:
```
https://github.com/muschellij2/ITKR/releases/download/v0.4.12.4/ITKR_0.4.12.4_R_x86_64-pc-linux-gnu.tar.gz
https://github.com/muschellij2/ANTsRCore/releases/download/v0.4.2.1/ANTsRCore_0.4.2.1_R_x86_64-pc-linux-gnu.tar.gz
https://github.com/muschellij2/ANTsR/releases/download/v0.6.2/ANTsR_0.6.2_R_x86_64-pc-linux-gnu.tar.gz
```


### Installing ichseg

The main R package that does the ICH segmentation in CT is `ichseg`:
https://github.com/muschellij2/ichseg.  After the 3 packages above are installed, you are ready to install the main pacakge `ichseg`
```{r}
devtools::install_github("muschellij2/extrantsr", upgrade_dependencies = FALSE)
devtools::install_github("muschellij2/ichseg", upgrade_dependencies = FALSE)
```


# Workflow 

Here we will have some data (DICOM format), that is unsorted and there are multiple pieces of data in there (such MRI scans, localizer scans, CTA, etc.).

## Sorting DICOM data (not solved)

Have a folder of DICOM data.  There can be multiple images in there, they will be sorted in the following steps.

We use the `tractor.base::sortDicomDirectories` function.  We need at least a specific version for sorting.  

```{r}
if (!("tractor.base" %in% installed.packages())) {
  install.packages("tractor.base")
}
tractor_version = packageVersion("tractor.base")
if (compareVersion(as.character(tractor_version), "3.1.3") < 0) {
  devtools::install_github(
    "tractor/tractor", 
    subdir = "tractor.base")
}
```

Now that you have the package installed, you should run the following steps for DICOM sorting (where you replace `"/path/to/dicom/files"` with the relevant directory):

```{r}
dicom_directory = "/path/to/dicom/files"
before_run = list.dirs(dir, recursive = FALSE)

# find all zip files - uncompress them, then delete zip files
all_zip = list.files(
  path = dir,
  pattern = "[.]zip$",
  recursive = TRUE, full.names = TRUE)
if (length(all_zip) > 0) {
  file.remove(all_zip)
}

all_zip = list.files(
  path = dir,
  pattern = "[.]rar$",
  recursive = TRUE, full.names = TRUE)
if (length(all_zip) > 0) {
  file.remove(all_zip)
}

# sort the data
res = tractor.base::sortDicomDirectories(
  directories = dicom_directory, 
  deleteOriginals = TRUE,
  ignoreTransferSyntax = TRUE
  )

# remove old directories
after_run = list.dirs(dicom_directory, recursive = FALSE)
new_dirs = setdiff(after_run, before_run)
old_dirs = intersect(after_run, before_run)

unlink(old_dirs, recursive = TRUE)
```

All files with the ending `.zip` will be deleted (sometimes they are duplicated).  If you want to keep these, I recommend using the `utils::unzip` command in R previous to running this.  The data will be copies, sorted, and the old data will be deleted.  

The structure of the directory specified in `dicom_directory` will be sorted based on Series (by default), based on Series unique ID (UID) based on DICOM tag 0x0020,0x000e by default.

### Subsetting: Not completed 

Now that the dat has been sorted, the relevant data can be subset.   The data for the PItCHPERFECT model requires the data be non-contrast CT data.  This means removing anything of the imaging modality MR (MRIs), CTAs (CT angiograms), and a slew of derived images (such as screen saves, dose reports, localizers, and 3D reconstructions).  

These can be subset using the DICOM header information:

* `Modality`: (0008,0060) tag
* `ImageType`: (0008,0008) tag  
* `Frame Type`: (0008,9007) tag
* `ConvolutionKernel`: (0018,1210) tag (Required if Frame Type (0008,9007) Value 1 of this frame is ORIGINAL. May be present otherwise.)
* `Convolution Kernel Group`: (0018,9316) tag
* `X-ray Tube Current`:  (0018,1151) tag X-ray Tube Current in mA.	
* `Exposure Time`: (0018,1150) tag Time of x-ray exposure in msec	

With this information, we will start removing unnecessary series.  We will use the `dcmtk` package for this:

```{r}
if (!("dcmtk" %in% installed.packages())) {
  devtools::install_github("muschellij2/dcmtk")
} else {
  dcmtk_ver = packageVersion("dcmtk")
  if (dcmtk_ver < "0.5.5") {
    devtools::install_github("muschellij2/dcmtk")
  }  
}
library(dcmtk)
```

We are reading in all the header information from each DICOM using the `dcmtk::read_dicom_header` function:
```{r}
n_dirs = length(new_dirs)
all_data = vector(mode = "list", length = n_dirs)
for (i in seq(n_dirs)) {
  basedir = new_dirs[i]
  hdr = dcmtk::read_dicom_header(file = paste0(basedir, "/*"))
  hdr$dir = basedir
  all_data[[i]] = hdr
}
```

NB: this data contains all the header information, not just those fields specified above, including protected health information (PHI).

```{r}
library(dplyr)
all_hdr = dplyr::bind_rows(all_data)
keep_tags = c("(0008,0008)", "(0008,0060)", "(0018,1210)",
              "(0018,1160)", "(0018,1151)", "(0018,0081)",
              "(0018,1150)", "(0018,0080)", "(0008,9007)",
              "(0018,9316)")
sub_hdr = all_hdr %>% 
  filter(tag %in% keep_tags) %>% 
  select(file, name, value)
```

## Converting DICOM to NIfTI data

Once we have a directory of DICOM files, we can convert them using to NIfTI using the DICOM to NIfTI converter [dcm2nii](https://www.nitrc.org/projects/dcm2nii/).  We use this through the `dcm2niir` package

The current workflow is to convert a directory (`directory_of_DICOMS`):

```{r, eval = FALSE}
library(dcm2niir)

out = dcm2nii(basedir = directory_of_DICOMS)
res_file = check_dcm2nii(out)
```

## Ensuring HU scale

We then read in the file, make sure it's within the standard range of HU and then write it out.


```{r, eval = FALSE}
library(neurobase)
####################################  
# window the image
####################################
window = c(-1024, 3071)
img = readnii(res_file)
img = window_img(img, window = window)
img = cal_img(img)
scl_slope(img) = 1
scl_inter(img) = 0
aux_file(img) = ""
descrip(img) = ""
writenii(img, filename = res_file)
````

## ICH Segmentation

This file can be passed into `ichseg::ich_segment`.  

```{r, eval = FALSE}
results = ichseg::ich_segment(res_file)
```


### Resampling to 1x1x1

In order to keep the dimensions of voxels the same, we can rigidly register to a template (which is done in `ich_segment`).

You can also resample the image to a 1x1x1$mm$ image:

```{r}
library(ANTsRCore)
img = antsImageRead(res_file)
res_img = resampleImage(img, resampleParams = c(1,1,1),
  useVoxels = FALSE)
```

The images should be on the same grid but not registered and not necessarily oriented the same way



